---
title: "lab1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create your first data frame

```{r cars}
animals <- c('Snake','Ostrich','Cat','Spider')
num_legs <- c(0,2,4,8)
animals_df <- data.frame(animals,num_legs)
animals_df
```

## Matrix operations

```{r pressure, echo=FALSE}
x_vector <- seq(from=12,to=2,by=-2)
x_vector
x_matrix <- matrix(x_vector,2,3)
x_matrix
y_matrix <- matrix(seq(1,4),2,2)
y_matrix
z_matrix <- matrix(seq(from=4,to=10,by=+2),2,2)
z_matrix
y_matrix_t <-t(y_matrix)
y_matrix_t
z_matrix_t <-t(z_matrix)
z_matrix_t
sum_matrix <- y_matrix + z_matrix
sum_matrix
sum_matrix_t <- y_matrix_t + z_matrix_t
sum_matrix_t
multi_matrix_one <- y_matrix%*%z_matrix
multi_matrix_two <- z_matrix%*%y_matrix
multi_matrix_one
multi_matrix_two
potmulti_matrix_one <- y_matrix*z_matrix
potmulti_matrix_two <- z_matrix*y_matrix
potmulti_matrix_one
potmulti_matrix_two
multi_matrix_three <- y_matrix%*%x_matrix
multi_matrix_three
```
The error mesage in XY is non-conformable arguments
The reasons behind this is the cols of the first matrix must equal to the rows of second matrix in matrix multiplication
```{r}
multi_matrix_four <- x_matrix%*%y_matrix
```
When a matrix multi the inverse of itself, it becomes a unit matrix
This generic function solves the equation a %*% x = b for x, where b can be either a vector or a matrix.
Usage
solve(a, b, ...)

```{r}
inverse_y_matrix <-solve(y_matrix)
inverse_y_matrix
mulit_y <-inverse_y_matrix%*%y_matrix
mulit_y
multi_x <-inverse_y_matrix%*%x_matrix
multi_x
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
