---
title: "lab2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Visualisation

```{r}
hawksSmall<-drop_na(select(Hawks,Age,Day,Month,Year,CaptureTime,Species,Wing,Weight,Tail))
dim(hawksSmall)
head(hawksSmall)
```

## Generate a histogram(one variables)
```{r}
weight_count_plot <- ggplot(data=hawksSmall,aes(x=Weight))+xlab('Weight(g')+ylab('Count')+geom_histogram(binwidth = 100)
weight_count_plot
```

## Generate a density plot
```{r}
tail_density_1 <- ggplot(data = hawksSmall,aes(x=Tail))+xlab('Tail (mm)')+ylab('Density')+geom_density(adjust=0.5)
tail_density_1
tail_density_2 <-ggplot(data = hawksSmall,aes(x=Tail))+xlab('Tail (mm)')+ylab('Density')+geom_density(adjust=1)
tail_density_1
```

## Bivariate density plot
```{r}
Bi_den_plot <- ggplot(data=hawksSmall,aes(x=Tail,color=Species))+xlab('Tail(mm)')+ylab('Density')+geom_density()+theme_bw()
Bi_den_plot
```

## violin plot
```{r}
violin_plot <- ggplot(data = hawksSmall,aes(x=Tail,fill=Species,y=Species))+xlab('Tail(mm)')+ylab('Species')+geom_violin()
violin_plot
```

## Scatter Plot
```{r}
scatter_plot <- ggplot(data=hawksSmall,aes(x=Tail,y=Weight,shape=Species,color=Species))+xlab('Tail(mm)')+ylab('Weight (gm)')+geom_point()
scatter_plot
```

## Trend lines and Facet Wraps
```{r}
facet_trendline <- ggplot(data=hawksSmall,aes(x=Tail,y=Weight,color=Species))+xlab('Tail(mm)')+ylab('Weight(gm)')+geom_point()+facet_wrap(~Species)+geom_smooth(method = 'lm')
facet_trendline
```

# Data wrangling

## Combine select with filter
columns are variable of example
use count function to count the total case in the dataFrame
```{r}
hSF <- hawksSmall %>% filter(Species=='RT'& Weight > 1000) %>% select(Wing,Weight,Tail)
hSF
count(hSF)
```
## Arrange function (sort the dataframe)
```{r}
arrange_hSF <- hSF %>% arrange(Wing)
head(arrange_hSF)
```
##Join and rename function
1.create a dataframe of full species name
unique() returns a vector, data frame or array with duplicate elements/rows removed
```{r}
help(unique)
species_code <- unique(hawksSmall$Species)
species_name_full <-c('Red-tailed',"Cooper's",'Sharp-shinned')
species_name_df <- data.frame(species_code,species_name_full)
species_name_df
```
2. fuse this two dataFrame with join function
```{r}
new_hawks <- species_name_df %>% rename(Species=species_code)
new_hawks
hawksFullNames <- Hawks %>% left_join(new_hawks) 
hawksFullName <- hawksFullNames %>% select(-Species) %>% rename(Species = species_name_full)
hawksFullName

```

## the mutate function (create a new columns)
```{r}
head(Hawks)
hawksWithBMI <- Hawks %>% mutate(bird_BMI = 1000*Weight/Wing^2) %>% select(Species,bird_BMI) %>% arrange(desc(bird_BMI))
hawksWithBMI
healthy_hawks <- hawksWithBMI %>% filter(bird_BMI < 100) %>% ggplot(aes(x=bird_BMI,y=Species,fill=Species))+ xlab('bird_BMI')+ ylab('Species')+geom_violin()+xlim(c(0,17))
healthy_hawks
```

##Summarize and groupby function(computes vector functions across the entire dataframe)

```{r}
hawksFullName %>%group_by(Species)%>% summarise(num_rows = n(),mn_wings=mean(Wing,na.rm = 1),md_wing=median(Wing,na.rm = 1),t_mn_wing=mean(Wing,na.rm = 1,trim = 0.01),tail_wing_ratio=mean(Tail/Wing,na.rm = 1))

```

## The across function (summarize to all columns at once)
e.g find out all the missing value by summarize and across
```{r}
hawksFullName %>% group_by(Species) %>% summarize(across(everything(),~sum(is.na(.x))))
```

# Exploratory data analysis

## Combining location estimators with the summarise function
```{r}
Hawks %>% summarise(Wing_mean= mean(Wing,na.rm=1),Wing_t_mean=mean(Wing,na.rm=1,trim=0.1),Wing_med=median(Wing,na.rm=1),Weight_mean=mean(Weight,na.rm=1),Weight_t_mean=mean(Weight,na.rm=1,trim=0.1),Weigth_med=median(Weight,na.rm=1))
```

## Combine group by function
```{r}
Hawks %>% group_by(Species) %>% summarise(Wing_mean= mean(Wing,na.rm=1),Wing_t_mean=mean(Wing,na.rm=1,trim=0.1),Wing_med=median(Wing,na.rm=1),Weight_mean=mean(Weight,na.rm=1),Weight_t_mean=mean(Weight,na.rm=1,trim=0.1),Weigth_med=median(Weight,na.rm=1))
```

## Robustness of location estimators
```{r}
hal<-Hawks$Hallux # Extract the vector of hallux lengths
hal<-hal[!is.na(hal)] # Remove any nans
```

To investigate the effect of outliers on estimates of location we generate a new vector called “corrupted_hall”
with 10 outliers each of value 100 created as follows:
```{r}
outlier_val<-100
num_outliers<-10
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
mean(hal)
mean(corrupted_hal)
```

investigate what happens as the number of outliers changes from 0 to 1000.
mean_vector
rep replicates the values in x.
```{r}
num_outliers_vect<-seq(0,1000)
means_vect<-c()
for(num_outliers in num_outliers_vect){
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
means_vect<-c(means_vect,mean(corrupted_hal))
}

```

median_vector
```{r}
medians_vect<-c()
for(num_outliers in num_outliers_vect){
  corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
  medians_vect<-c(medians_vect,median(corrupted_hal))
}
```

t_means_vector
```{r}
t_means_vect <-c()
for(num_outliers in num_outliers_vect){
  corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
  t_means_vect<-c(t_means_vect,mean(corrupted_hal,trim=0.1))
}

```

Combine these vectors into a data frame 
```{r}
df_means_medians<-data.frame(num_outliers=num_outliers_vect,
mean=means_vect,t_mean=t_means_vect,
median=medians_vect)

```

use the code below to reshape and plot the data
```{r}
df_means_medians%>%
pivot_longer(!num_outliers, names_to = "Estimator", values_to = "Value")%>%
ggplot(aes(x=num_outliers,color=Estimator,
linetype=Estimator,y=Value))+
geom_line()+xlab("Number of outliers")
```

## Box plot and outliers
```{r}
Hawks %>% ggplot(aes(x=Species,y=Weight))+xlab('Species')+ylab('Weight')+geom_boxplot()
```

count the number of outliers
```{r}
numberof_outliers <- function(species){
  q25<-quantile(species,0.25,na.rm=1)
  q75<-quantile(species,0.75,na.rm=1)
  iq_range<- q75-q25
  outliers <- species[(species>q75+1.5*iq_range) | (species<q25-1.5*iq_range)]
  return(length(outliers))
}

CH <- Hawks %>% filter(Species=='CH')
RT <- Hawks %>% filter(Species=='RT')
SS <- Hawks %>% filter(Species=='SS')
CH_outliers<- numberof_outliers(CH$Weight)
RT_outliers<- numberof_outliers(RT$Weight)
SS_outliers<- numberof_outliers(SS$Weight)
Speices<-c('CH','RT','SS')
num_outliers_weight<-c(CH_outliers,RT_outliers,SS_outliers)
outliers_df <- data.frame(Speices,num_outliers_weight)
outliers_df

```

