---
title: "lab3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Missing data and iteration
## impute by median
```{r}
impute_by_median<-function(x){
  mu<-median(x,na.rm=1)
  impute_f<-function(z){
    if(is.na(z)){
      return(mu)
    }else{
      return(z)
    }
  }
  return(map_dbl(x,impute_f))
}


v<-c(1,2,NA,4)
impute_by_median(v)
```
## map2 function
iterates over two variables in paralled rather than one

```{r}
x<-seq(0,10,0.1)
y<-5*x+1
df_xy <- data.frame(x,y)
df_xy %>% head(5)
df_xy %>% mutate(z=map2_dbl(x,y,~.x+.y))%>%
  head(5)
```

```{r}
 sometimes_missing <- function(index,value){
   if(index%%5==0){
     return(NA)
   }else{
     return(value)
   }
 }
sometimes_missing(14,25)
sometimes_missing(15,25)
```
```{r}
new_y <- map2_dbl(row_number(y),y,sometimes_missing)
df_xy_missing <- data.frame(x,y=new_y)
df_xy_missing
```

```{r}
df_xy_impute <- df_xy_missing %>% mutate(x=x,y=impute_by_median(y))
df_xy_impute                                     
                                          
```
Combine the dataframes df_xy, df_xy_missing and df_xy_impute and plot it
```{r}
df_xy <- df_xy %>%mutate(source='original')
df_xy_missing <- df_xy_missing %>% mutate(source='corrupted')
df_xy_impute <- df_xy_impute %>% mutate(source='imputed')
df_combined <- rbind(df_xy,df_xy_missing,df_xy_impute)

df_combined %>% ggplot(aes(x=x,y=y,color=source))+geom_point()+facet_wrap(~source)+geom_smooth(method='lm')

```
 
# Tidying data with pivot function
1. read xls by readxl
```{r}
folder_path <- 'C:/Users/Cam/Desktop/firstRproject/firstRProject/'
file_name <- 'HockeyLeague.xlsx'
file_path <- paste(folder_path,file_name,sep = "")
wins_data_frame <- read_excel(file_path,sheet = 'Wins')
```
 
```{r}
wins_data_frame %>% select(1:5) %>% head(3)
rename_data <- wins_data_frame %>% rename(Team=...1)
longer_data <- rename_data %>% pivot_longer(cols = !Team, names_to = 'Year',values_to = 'Value',)
separate_data <- longer_data %>% separate(Value, into = c('Wins','Total'), sep='of', convert = TRUE)
wins_tidy <- separate_data %>% mutate(Team = as.character(Team),Year= as.integer(Year), Wins = as.integer(Wins), Total = as.integer(Total))
wins_tidy%>% dim()
wins_tidy %>% head(5)
```
```{r}
losses_data <- read_excel(file_path,sheet = 'Losses')
losses_tidy <- losses_data %>% rename(Team=...1) %>% pivot_longer(cols = !Team,names_to = 'Year',values_to = 'Value') %>% separate(Value,into = c('Losses','Total'), sep = 'of', convert = TRUE) %>% mutate(Team = as.character(Team),Year=as.integer(Year),Losses=as.integer(Losses),Total=as.integer(Total))

```
```{r}
data_combine <- inner_join(wins_tidy,losses_tidy)
hockey_df <- data_combine %>% mutate(Wins=Wins,Losses=Losses,Total=Total,Draws=Total-(Wins+Losses),Wins_rt=Wins/Total,Losses_rt=Losses/Total,Draws_rt=Draws/Total)
head(hockey_df)

```
## conclude

```{r}
hockey_conclude <- hockey_df %>% group_by(Team) %>% select(Wins_rt,Losses_rt,Draws_rt) %>%summarise(win_mean=mean(Wins_rt,na.rm=TRUE),lose_mean=mean(Losses_rt,na.rm=TRUE),draw_mean=mean(Draws_rt,na.rm=TRUE),win_median=median(Wins_rt,na.rm=TRUE),lose_median=median(Losses_rt,na.rm=TRUE),draw_median=median(Draws_rt,na.rm=TRUE))
hockey_conclude %>% arrange(win_mean)

```

## Most correlated variables
```{r}
head(penguins)
penguins_data <- penguins %>% select(-species,-island,-sex)
penguins_data%>% select(all_of(body_mass_g))
c(names(penguins_data))



```
## correlate function
-unlist simplifies it to produce a vector which contains all the atomic components which occur in x.
-abs(x) computes the absolute value of x
-which() Give the TRUE indices of a logical object, allowing for array indices.

```{r}
data_name <- c(names(penguins_data))
year<-c('1','2','3')
top_correlates_by_var <- function(df){
 
  }
  
}
top_correlates_by_var(penguins_data)
```


