---
title: "lab4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bayes theorem
```{r}
p_A<-0.9
p_B_given_A<-0.8
p_not_B_given_not_A<-0.75
p_B<-p_B_given_A*p_A+(1-p_not_B_given_not_A)*(1-p_A)
p_A_given_B<-p_B_given_A*p_A/p_B ##Bayes theorem
p_A_given_B
```

## Conditional probabilities
Suppose that A,B ∈ E and A ⊆ B. Give an expression for P(A|B) in terms of P(A) and P(B). What
about when P(A \ B) = 0?
If P(A \ B) = 0 then P(A) = P(A ∩ B) + P(A ∩ Bc) = P(A ∩ B) + P(A \ B) = P(A ∩ B). Hence
P(A|B) = P(A ∩ B) P(B) = P(A) P(B).
Note that this includes the case where A ⊆ B so A \ B = ∅

Suppose that A, B ∈ E with A∩B = ∅. Give an expression for P(A|B). What about when P(A∩B) = 0?
If P(A ∩ B) = 0 then P(A|B) = P(A ∩ B)/P(B) = 0. This includes the case where A ∩ B = ∅.

Suppose that A, B ∈ E with B ⊆ A. Give an expression for P(A|B). What about when P(B \ A) = 0?
If P(B \ A) = 0 then P(B) = P(A ∩ B) + P(Ac ∩ B) = P(A ∩ B) + P(B \ A) = P(A ∩ B) and so
P(A|B) = P(A ∩ B)/P(B) = P(B)/P(B) = 1. This includes the case where B ⊆ A, so B \ A = ∅.

Suppose that A ∈ E. Give an expression for P(A|Ω) in terms of P(A). 
In this case we have A ⊆ Ω and P(Ω) = 1, so P(A|Ω) = P(A)/P(Ω) = P(A)

Show that given three events A, B, C ∈ E we have P(A ∩ B ∩ C) = P(A|B ∩ C) · P(B|C) · P(C).
We have P(A|B ∩ C) = P(A ∩ B ∩ C)/P(B ∩ C) and P(B|C) = P(B ∩ C)/P(C), so P(A|B ∩ C) · P(B|C) · P(C) = P(A ∩ B ∩ C) P(B ∩ C) · P(B ∩ C) P(C) · P(C) = P(A ∩ B ∩ C)

Show that given three events A, B, C ∈ E we have P(A|B ∩ C) = P(B|A∩C)·P(A|C) P(B|C) .
We substitute in the definition of each of the constituting conditional probabilities to see that P(B|A ∩ C) · P(A|C) P(B|C) =  P(A ∩ B ∩ C) P(A ∩ C)  ·  P(A ∩ C) P(C)  ·  P(B ∩ C) P(C)  −1 = P(A ∩ B ∩ C) P(B ∩ C) = P(A|B ∩ C), as required.

## Sampling with replacement
compute the number of different subsets of size 3 from a collectiom of size 8
choose()
example(10)
```{r cars}
choose(8,3)

prob_red_spheres <- function(z){
  possible_subsets <- choose(35,z)
  poss_red <- (3/10)^z
  poss_not_red <- (7/10)^(35-z)
  prob <- possible_subsets*poss_red*poss_not_red
  return(prob)
}

prob_red_spheres(20)

```

prob red df
```{r}
num_reds <- seq(1,35)
prob <- prob_red_spheres(num_reds)
prob_by_num_reds <- data.frame(num_reds,prob)
prob_by_num_reds %>% head(3)
```

prob red graph
```{r}
red_line_graph <- prob_by_num_reds %>% ggplot(aes(x=num_reds,y=prob))+geom_line()+xlab('Number of reds')+ylab('Probability')
red_line_graph
```

## Sample and set seed
sample takes a sample of the specified size from the elements of x using either with or without replacement.
Get a different answer every time we run the code is problematic
To avoid this process we set a random seed set.seed() to get the same output every time
```{r}
sample(10,35,replace=TRUE)


for(i in 1:5){
  print(sample(100,5,replace=FALSE))
  #result diff
}

for(i in 1:5){
  set.seed(0)
  print(sample(100,5,replace=FALSE))
  #result same because of set.seed()
}
```

## sampling_with_replacement_simulation
trial=test
```{r}
num_trials <- 1000 #set the number of trials
set.seed(0) #set the random seed
##sample(x, size, replace = FALSE, prob = NULL)
##that sample(x) generates a random permutation of the elements of x (or 1:x).
sampling_with_replacement_simulation <- data.frame(trial=1:num_trials) %>% mutate(sample_balls=map(.x=trial,~sample(10,35,replace=TRUE)))

#count red ball, 1,2,3 = red
sampling_with_replacement_simulation <-sampling_with_replacement_simulation %>% mutate(num_reds=map_dbl(.x=sample_balls,~sum(.x<4)))


#pull() is similar to $. sample$num_reds
num_reds_in_simulation <- sampling_with_replacement_simulation %>% pull(num_reds)

### add a column which gives the number of trials which gave each number of reds
prob_by_num_reds <- prob_by_num_reds %>% mutate(simulation_count=map_dbl(.x=num_reds,~sum(num_reds_in_simulation==.x)))

##expected count
### add a column which gives the expected number of reds based on the probability formula
prob_by_num_reds <- prob_by_num_reds %>% mutate(expected_count = num_trials*prob)

prob_by_num_reds %>% rename(Simulation = simulation_count,Expected=expected_count)%>% pivot_longer(cols=c('Simulation','Expected'),names_to = 'Type',values_to = 'count') %>% ggplot(aes(num_reds,count)) + geom_line(aes(linetype=Type,color=Type)) + scale_linetype_manual(values = c('solid','dashed')) + theme_bw() + xlab('Numbers of reds') + ylab('Count')
                  


```
## Sample without replacement
```{r}
##Try this initially with a small number of simulations. Increase your number of simulations to about a relatively large number to get a more accurate answer, once everything seems to be working well
##1. First set a random seed;
set.seed(0) 
##2. Next set a number of trials, a number of reds, a number of blues, a number of greens and a sample size;
num_trials<- 1000
n_red<-50
n_blue<-30
n_green<-20
sample_size<-10
##3. Now use a combination of the functions sample(), mutate() and map() to generate your samples. Here you are creating sample of size 10 from a collection of 100 balls - the sampling is done without replacement;
total<-n_red+n_blue+n_green
sampling_without_replacement_simulation<-data.frame(trial=1:num_trials)%>%
mutate(sample_balls=map(.x=trial,~sample(total,sample_size,replace = FALSE)))%>%
# simulate without replacement the sampling of 10 balls from a bag of 100
mutate(num_reds=map_dbl(.x=sample_balls,~sum(.x<=n_red)))%>%
mutate(num_blues=map_dbl(.x=sample_balls,~sum((n_red<.x)&(.x<=n_red+n_blue))))%>%
mutate(num_greens=map_dbl(.x=sample_balls,~sum((n_red+n_blue<.x)&
(.x<=n_red+n_blue+n_green))))%>%
# count the number of each colour the minimum of the three counts
mutate(colour_missing=(pmin(num_reds,num_blues,num_greens)==0))
# look for missing colours
sampling_without_replacement_simulation
prob_missing_simulation<-mean(sampling_without_replacement_simulation%>%
pull(colour_missing))

```

Once you have a simulation based answer you can try and use “combinations” with n choose r to compute the
probability directly. First aim and compute the number of subsets of size 10 from 100 which either entirely
miss out one of the subsets Reds = {1, . . . , 50}, Blues = {51, . . . , 80}, Greens = {81, . . . , 100}. Be careful
not to double count some of these subsets! Once you have computed all such subsets combine with the
formula for the total number of subsets of size 10 from a set of 100, to compute the probability.

We first compute the number of subsets of size 10 containing no reds, no blues and no greens:
1. NR - The number of subsets with no reds is 50 c 10
2. NB 70 c 10
3. NG 80 c 10
Next we count the number of all red, all blue and all green:
1. R 50 c 10
2. B 30 c 10
3. G 20 c 10
Now the total number of subsets with a missing colour is Q = (NR + NB + NG) − (MR + MB + MG).
The total number of subsets of size 10 from a collection of a 100 is T = 100 c 10
```{r}
NR=choose(total-n_red,sample_size)
NB=choose(total-n_blue,sample_size)
NG=choose(total-n_green,sample_size)
MR=choose(n_red,sample_size)
MB=choose(n_blue,sample_size)
MG=choose(n_green,sample_size)
Q=NR+NB+NG-MR-MB-MG
TS=choose(total,sample_size)
prob_missing_theory=Q/TS
prob_missing_simulation
prob_missing_theory
```

## Mutual independence and pair-wise independent
Verify that P(A ∩ B) = P(A) · P(B), P(A ∩ C) = P(A) · P(C) and P(B ∩ C) = P(B) · P(C). Hence, we deduce that the events A, B, C are pair-wise independent.
Observe that A ∩ B = {(1, 1, 0)} so P(A ∩ B) = 14 = P(A) · P(B). We proceed in the same wau for the remaining cases.

What is A ∩ B ∩ C? What is P(A ∩ B ∩ C)?
Observe that A ∩ B ∩ C = ∅ and so P(A ∩ B ∩ C) = 0

Are the events A,B,C mutually independent?
We have P(A∩B∩C) = 0 ̸= 18 = P(A)·P(B)·P(C). Hence, these events cannot be mutually independen