---
title: "lab7_markdown"
output: html_document
---

```{r}
library(PropCIs) ## wilson method
library(Stat2Data)
library(tidyverse)
library(palmerpenguins)
library(PairedData)
library(boot) ##bootstrap method
library(usethis)
data("Hawks")
data('Barley')
data('Airlines')
data('penguins')
use_git_config(user.name = "chaokinhong", user.email = "3484715033@qq.com")
```

#Student's t-confidence interval
sd and mean function must input a vector
library(Stat2Data)
data(Hawks)
```{r}
red_tail_hawks <- c(Hawks %>% filter(Species=='RT') %>% pull(Weight))
red_tail_hawks_data <- Hawks %>% filter(Species == 'RT') 
head(red_tail_hawks_data)
alpha <- 0.05
n <- length(red_tail_hawks)
sample_sd <- sd(red_tail_hawks,na.rm=1)
sample_mean <- mean(red_tail_hawks,na.rm = 1)
## 99% confidence level
t <- qt(1-alpha/2, df=n-1)

confidence_interval_l <- sample_mean-t*sample_sd/sqrt(n)
confidence_interval_u <- sample_mean+t*sample_sd/sqrt(n)
confidence_interval <- c(confidence_interval_l,confidence_interval_u)
confidence_interval

```
## Assumption (Proof it is a Gaussian distribution)
```{r}
## density plot
ggplot(data=red_tail_hawks_data,aes(x=Weight)) + geom_density()+theme_bw()+xlab('Weight')
## qq plot
## if qq-plot point close to the straight line, it is a Gaussian distribution
ggplot(data=red_tail_hawks_data,aes(sample=Weight))+theme_bw()+stat_qq()+stat_qq_line(color='blue')
```

## One sample t-test
t.test function must input a vector
t.test(x, y = NULL,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95, ...)

U0: mu =40
U1: mu != 40
library(palmerpenguins)
data(penguins)
```{r}
bill_adelie <- c(penguins %>% filter(species=='Adelie') %>% pull(bill_length_mm))
t.test(x=bill_adelie,mu=40,conf.level = 0.99)

```

## Inplementing a t-test
u=u0
```{r}
t_test_in <- function(sample,u){
  sample_size <- length(sample)
  sample_mean <- mean(c(sample),na.rm = 1)
  sample_sd <- sd(c(sample),na.rm = 1)
  t_value <- (sample_mean-u)/(sample_sd/sqrt(sample_size))
  p_value <- 2*(1-pt(abs(t_value),df=sample_size-1))
  return(p_value)
}
t_test_in(bill_adelie,40)
```

## Paired t-test 
effect_size (0:1)
library(PairedData)
data('Barley')
```{r}
t.test(x=Barley$Glabron,y=Barley$Velvet,paired = TRUE)
diff <- c(Barley$Glabron- Barley$Velvet)
mean_dif <- mean(diff,na.rm = 1)
s_dif <- sd(diff,na.rm = 1)
effect_size <- mean_dif/s_dif
effect.size
```
## Implement confidence interval
```{r}
student_t_confidence_interval<-function(sample,confidence_level){
sample<-sample[!is.na(sample)] # remove any missing values
n<-length(sample) # compute sample size
mu_est<-mean(sample) # compute sample mean
sig_est<-sd(sample) # compute sample sd
alpha = 1-confidence_level # alpha from gamma
t<-qt(1-alpha/2,df=n-1) # get student t quantile
l=mu_est-(t/sqrt(n))*sig_est # lower
u=mu_est+(t/sqrt(n))*sig_est # upper
return(c(l,u))
}

```

## Coverage simulation
```{r}
num_trails <- 10000
sample_size <- 30
mu_0 <- 1
sigma_0 <- 3
alpha <- 0.05
set.seed(0)
single_alpha_coverage_simulation_df <- data.frame(trial=seq(num_trails)) %>% mutate(sample=map(.x=trial,.f=~rnorm(n=sample_size,mean=mu_0,sd=sigma_0))) %>% mutate(ci_interval=map(.x=sample,.f=~student_t_confidence_interval(.x,1-alpha))) %>% mutate(cover=map_lgl(.x=ci_interval,.f=~((min(.x)<=mu_0)&(max(.x)>=mu_0)))) %>% mutate(ci_length = map_dbl(.x=ci_interval,.f=~(max(.x)-min(.x))))

single_alpha_coverage_simulation_df %>% pull(cover) %>% mean()
```
## Wilson's confidence interval
library(PropCIs)
```{r}

red_tail_hawks <- Hawks %>% filter(Species=='RT') %>% pull(Weight)
mean(red_tail_hawks,na.rm = 1)
alpha <- 0.01
Weight_more_than_1000 <-  Hawks %>% filter(Species=='RT' & Weight > 1000 ) 
num_lrg <- count(Weight_more_than_1000)$n
sample_size <- length(red_tail_hawks)
scoreci(x=num_lrg,n=sample_size,conf.level = 1-alpha)
```

## The Binomial test
```{r}
head(Airlines)
on_time <- count(Airlines %>% filter(OnTime=='yes'))$n
sample_size <- count(Airlines)$n
binom.test(x=on_time,n=sample_size,p=875/1000,alternative = 'two.sided')

```
## Bootstrap confidence intervals
```{r}
hawks_weight <- Hawks %>% pull(Weight)
set.seed(123)  ## set random seed

compute_median <- function(df,idx,col_name){
  sub_sample <- df %>% slice(idx) %>% pull(all_of(col_name))
  return(median(sub_sample,na.rm = 1))
}

##generate bootstrap statistics
boot_result <- boot(data= Hawks,statistic = compute_median,col_name='Weight',R=10000)

boot_result
## compute 99% interval
boot.ci(boot.out = boot_result,type = 'basic',conf = 0.99)

```

